---
title: "153_project"
output:
  pdf_document: default
  html_document: default
date: '2022-04-05'
---

```{r}
library(tidyverse)
library(TSA)
library(forecast)
library(astsa)
source('cleaning.R')
```

  
# Original data
```{r}
ozone <- phoenix$o3
ozone = ozone - mean(ozone) # mean centered
plot(ozone,type ="l")
acf(ozone,lag.max = 30)
pacf(ozone)
```
1. The original data have clear sign of seasonality, but there seems to be linear upward trend.
2. The ACF PACF plot shows the strong sign of seasonality

# Sinusoidal fitting
```{r}
t = 1:length(ozone)
# Check the periodogram
periodo = periodogram(ozone,plot=TRUE,ylab="Periodogram", xlab="Frequency") # There is one significant peak

# Get the high magnitudes in descending order
order_spec = sort(periodo$spec,decreasing = TRUE)

# Get the frequency that gives max magnitude
first_max = order_spec[1]
first_maximizing_freq = periodo$freq[periodo$spec==first_max]
first_sin_max = sin(2*pi*first_maximizing_freq*t)
first_cos_max = cos(2*pi*first_maximizing_freq*t)

# Max Sinusoidal fitting
ozone_sinusoid_model = lm(ozone ~ first_sin_max*(1+t)+first_cos_max*(1+t))
print(ozone_sinusoid_model$coefficients)

# Overlay the sinusoidal fitting over the original plot
plot(ozone,type = "l")
lines(t,ozone_sinusoid_model$fitted.values,col = "red")

# Get the residual, hoping for removing seasonality
ozone_sinusoid_residual = ozone_sinusoid_model$residuals
plot(ozone_sinusoid_residual,type = "l") # residual seems to be stationary
stats::acf(ozone_sinusoid_residual)
stats::pacf(ozone_sinusoid_residual)
```
Vt = ozone
(1+B)Vt = Xt
(Vt - f(t)) = Xt
$$f(t) = -0.003 + 0.000042*t - 0.0012*sin(t) -0.0123*cos(t) + 0.000002*t* sin(t) -0.000011*t*cos(t)$$ 

 1. There is one significant peak in the periodogram
 2. The residual plot shows some seasonality but the plot seems to be AR process and possible seasonal ARMA



# Residual Fitting : SARIMA(2,0,0)(1,0,1)12 
```{r}
model1 <- sarima(ozone_sinusoid_residual, p=2, d=0, q=0, P=1, D=0, Q=1, S=12) # fit the model

coeff_table <- as.data.frame(model1$ttable)
coeff_table <-coeff_table %>% mutate(ci_lower = Estimate-1.96*SE,ci_upper =  Estimate+1.96*SE) 
coeff_table # show estimated coefficient and its ci
```
# Residual Fitting : SARIMA(1,0,0)(1,0,1)12 
```{r}
model2 <- sarima(ozone_sinusoid_residual, p=1, d=0, q=0, P=1, D=0, Q=1, S=12) # fit the model

coeff_table <- as.data.frame(model2$ttable)
coeff_table <-coeff_table %>% mutate(ci_lower = Estimate-1.96*SE,ci_upper =  Estimate+1.96*SE) 
coeff_table # show estimated coefficient and its ci
```

# Residual Fitting : SARIMA(2,1,1)(1,0,1)12 
```{r}
model3 <- sarima(ozone_sinusoid_residual, p=2, d=1, q=1, P=1, D=0, Q=1, S=12) # fit the model

coeff_table <- as.data.frame(model3$ttable)
coeff_table <-coeff_table %>% mutate(ci_lower = Estimate-1.96*SE,ci_upper =  Estimate+1.96*SE) 
coeff_table # show estimated coefficient and its ci
```

# Residual Fitting : SARIMA(2,0,1)(1,0,1)12 
```{r}
model4 <- sarima(ozone_sinusoid_residual, p=2, d=0, q=1, P=1, D=0, Q=1, S=12) # fit the model

coeff_table <- as.data.frame(model4$ttable)
coeff_table <-coeff_table %>% mutate(ci_lower = Estimate-1.96*SE,ci_upper =  Estimate+1.96*SE) 
coeff_table # show estimated coefficient and its ci
```
# Evaluation Matrix
```{r}
# AIC, AICc, BIC
eval<- function(model){
  return (c(model$AIC, model$AICc,model$BIC))
}

m1_evaludation = eval(model1)
m2_evaludation = eval(model2)
m3_evaludation = eval(model3)
m4_evaludation = eval(model4)

eval_matrix = rbind(m1_evaludation,m2_evaludation,m3_evaludation,m4_evaludation)
rownames(eval_matrix) = c("SARIMA(2,0,0)(1,0,1)12","SARIMA(1,0,0)(1,0,1)12",
                          "SARIMA(2,1,1)(1,0,1)12 ","SARIMA(2,0,1)(1,0,1)12 ")
colnames(eval_matrix) = c("AIC","AICc","BIC")
eval_matrix
```
The first two models, SARIMA(2,0,0)(1,0,1)12 , SARIMA(1,0,0)(1,0,1)12 gives the lowest AIC,AICc,BIC


# Cross Validation 
  Train : 2004 ~ 2011
  Test : 2012 - 2015
```{r, warning=FALSE}
sse1 = c()
sse2 = c()
test_years = seq(10,15,1)
for (year in test_years) {

train_index = 1:(12*(year-4))
test_index = (12*(year-4)+1):(12*(year-4+1))

train <- ozone_sinusoid_residual[train_index]
test <- ozone_sinusoid_residual[test_index]
m1_forecast <- sarima.for(train, p=2, d=0, q=0, P=1, D=0, Q=1, S=12, n.ahead=12)$pred
m2_forecast <- sarima.for(train, p=1, d=0, q=0, P=1, D=0, Q=1, S=12, n.ahead=12)$pred

sse1 = c(sse1,sum((m1_forecast - test)^2))
sse2 = c(sse2,sum((m2_forecast - test)^2)) }

sse = rbind(sum(sse1),sum(sse2))
rownames(sse) = c("SARIMA(2,0,0)(1,0,1)12","SARIMA(1,0,0)(1,0,1)12")
colnames(sse) = c("SSE")
print(sse) 
```
The SSE for both model are low, but SARIMA(2,0,0)(1,0,1)12 is slighly better
 
 
 
# Future 10 values prediction
```{r}
m1_forecast <- sarima.for(ozone_sinusoid_residual, p=2, d=0, q=0, P=1, D=0, Q=1, S=12, n.ahead=10)$pred

ts = as.data.frame(145:154)
predict_t = 145:154
prediction = c()
coef = ozone_sinusoid_model$coefficients

for (t in predict_t){
  pred = coef[1]+coef[2]*sin(t) + coef[3]*t+ coef[4]*cos(t)*t+ coef[5]*sin(t)*t + coef[6]*cos(t)*t
  prediction = c(prediction,pred)
}


m1_forecast = as.vector(m1_forecast)
prediction = as.vector(prediction)
predictions = m1_forecast+prediction

ozone2 = c(ozone,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA)

ts2 = 1:154
preds = rep(NA, 144)
preds = c(preds,predictions)

plot(ozone2,type ="l")
lines(ts2,preds,col = "red")
```

