---
title: "153_project"
output: html_document
date: '2022-04-05'
---

```{r}
library(tidyverse)
library(TSA)
library(forecast)
library(astsa)
source('cleaning.R')

```

# Data Preprocessing
 - Get the index of the last day of the each year
```{r}
phoenix = phoenix %>% mutate(date = paste(year,paste(month,day,sep = "-"),sep = "-"))
years = seq(2000,2010,1)
year_index = c()
for (year in years){
  year_index = c(year_index,which(phoenix$date == paste(year,sep = "-", "12-31")))
}

phoenix[year_index,]
```
```{r}
paste("1",sep = "-","1")
```


# Original data
```{r}
ozone <- phoenix$o3[1:row_index_2010_12_31]

plot(ozone,type ="l")
acf(ozone)
pacf(ozone)
```
The original data have clear sign of seasonality, but there seems to be no trend.
# First difference
```{r}
diff_ozone= diff(ozone)
plot(diff_ozone,type= "l")
acf(diff_ozone)
pacf(diff_ozone)
```
The ACF of the first differenced data shows that there exists seasonality. Also, we can observe high autocorrelations at lag 1,2, and 3 it is reasonable to take the difference of order 3 or possibly 4. 
# Higer order difference
```{r}
diff_3_ozone= diff(diff(diff(ozone)))
diff_4_ozone = diff(diff(diff(diff(ozone))))
plot(diff_3_ozone,type = "l")
acf(diff_3_ozone)
pacf(diff_3_ozone)

plot(diff_4_ozone,type = "l")
acf(diff_4_ozone)
pacf(diff_4_ozone)
```
The ACF plot has some seasonality and pacf plot shows the sign of MA process
 - possible fitting: MA process, sinusoidal fitting

# thrice difference + sinusoidal fitting
```{r}
# Check the periodogram
periodo = periodogram(diff_4_ozone,plot=TRUE,ylab="Periodogram", xlab="Frequency")
# has two peaks and many leakage

# Get the high magnitudes in descending order
order_spec = sort(periodo$spec,decreasing = TRUE)

# Get the frequency that gives max magnitude
first_max = order_spec[1]
first_maximizing_freq = periodo$freq[periodo$spec==first_max]

# Get the frequency that gives second max magnitude
second_max = order_spec[2]
second_maximizing_freq = periodo$freq[periodo$spec==second_max]

t = 1:length(diff_4_ozone)

first_sin_max = sin(2*pi*first_maximizing_freq*t)
first_cos_max = cos(2*pi*first_maximizing_freq*t)
second_cos_max = cos(2*pi*second_maximizing_freq*t)
second_sin_max = cos(2*pi*second_maximizing_freq*t)

# Max Seasonal fitting
ozone_sinusoid_model = lm(diff_4_ozone ~ t+ first_sin_max+first_cos_max)

# Get the residual, hoping for removing seasonality
ozone_sinusoid_residual=ozone_sinusoid_model$residuals
plot(ozone_sinusoid_residual,type = "l") # residual seems to be stationary

acf(ozone_sinusoid_residual)
pacf(ozone_sinusoid_residual)
```
Looks like MA(1) model
# Fitting MA(1)
```{r}

```
# AIC, AICc, BIC
```{r}
auto.arima(ozone_sinusoid_residual)


ma1_model <- sarima(ozone_sinusoid_residual, p=0, d=0, q=1, P=0, D=0, Q=0, S=0) # fit the model

coeff_table <- as.data.frame(ma1_model$ttable)
coeff_table <-coeff_table %>% mutate(ci_lower = Estimate-1.96*SE,ci_upper =  Estimate+1.96*SE) 
coeff_table # show estimated coefficient and its ci



# AIC, AICc, BIC
eval<- function(model){
  return (c(model$AIC, model$AICc,model$BIC))
}
evaluation = eval(ma1_model)

```
# Cross Validation
```{r}
years = seq(6,10,1) # from 2006 to 2010
sse1 = c()
differenced_year_index = year_index-4  # differenced_year_index element represent
                                       #  index of the 200i 12 31, i = 0, 1,2,3...10

for (year in years) {

train_index = 1:differenced_year_index[year-1]
test_index = (differenced_year_index[year-1]+1):differenced_year_index[year]

train <- ozone_sinusoid_residual[train_index]
test <- ozone_sinusoid_residual[test_index]
              
ma1_forecast <- sarima.for(train, p=0, d=0, q=1, P=0, D=0, Q=0, S=0, n.ahead=length(test_index))$pred

sse1 = c(sse1,sum((ma1_forecast - test)^2))

}

sse1

```



